Optimised Cost Considering Huffman Code for Biological Data Compression

### 2015-01-30 Article/Article.tex
with mega

%% main text
\section{Introduction}
\label{sec1}
In the recent years, application of battery-powered portable devices, e.g. laptop computers and mobile phones has increased significantly. Proper representation of digital data and their transmission efficiency has become a primary concern for digital community because it affects the performance, reliability, and the cost of computation in both portable and non-portable devices. CMOS technologies were developed in order to reduce the power consumption both in data processing and transmission. In order to increase transmission speed and reduce transmission cost, parallel data transmission methods are widely used. However, parallel transmission is limited to short distance communications, e.g. locally connected devices, internal buses. Ruling out the possible availability of parallel transmission links over long distance, we are left with its serial alternative only. If we attempt to transfer big files, e.g. DNA sequences, over a serial transmission link then it would take a significant amount of time. However, we cannot overlook this problem because at present parallel processing is widely used to increase throughput and in parallel processing architecture, processing units are usually distributed in different physical locations and task sharing is a must in such architecture.

Data encoding techniques came into action to improve the data transmission efficiency over serial communication medium by compressing data before transmitting. Efficiency can be measured in terms of incurred cost, required storage space, consumed power, time spent and likewise. Data must be encoded to meet the purposes like: unambiguous retrieval of information, efficient storage, efficient transmission and etc. Let a message consist of sequences of characters taken from an alphabet $\Sigma$, where  $\alpha_1,\alpha_2,\alpha_3\ldots,\alpha_r$ are the elements that represent the characters in the source $\Sigma$. The length of $\alpha_i$ represents its cost or transmission time, i.e., $c\left(\alpha_i\right)= length(\alpha_i)$. A codeword $w_i$ is a string of characters in $\Sigma$, i.e., $w_i\in\Sigma^{+}$. If a codeword is $w_i=\alpha_{i1},\alpha_{i2},\ldots,\alpha_{in}$, then the length or cost of the codeword is the sum of the lengths of its constituent elements:

@@

\section{Results And Discussion}
\label{sec4}
The approach has been evaluated with different real genomic biological data, these genomes were downloaded from a recent version of The National Center for Biotechnology Information (NCBI) available on $(http://www.ncbi.nlm.nih.gov)$. We focused on the sequences alone, ignoring any header and any other exogenous information. In table () the different data set are described with the size of each of them and the references on the biological data bank.
\begin{table}
\small
\label{datas}
\caption{Dataset description}
\begin{center}
\begin{tabular}{c  c  c c}
\toprule
$\textbf{Data sets}$ &$\textbf{Name}$ &	$\textbf{Size (MegaOctet)}$ &	$\textbf{Reference}$ \\\hline
Genome 1& Mycobacterium smegmatis &6,988,302 & CP009496  \\\hline

Genome 2& Amycolatopsis benzoatilytica &  8,704,271& NZ\_KB912942 \\\hline

\multirow{2}{*}{Genome 3}&Mycobacterium rhodesiae& \multirow{2}{*}{6,415,739} & \multirow{2} {*}{CP003169}\\ 
&NBB3& &\\
\hline

\multirow{2}{*}{Genome 4 }&Streptomyces bottropensis& \multirow{2}{*}{8,955,726} &
\multirow{2}{*}{NZ\_KB911581} \\ 
& ATCC 25435 \\
\hline

\multirow{2}{*}{Genome 5}&Mycobacterium smegmatis& \multirow{2}{*}{6,988,269 } &\multirow{2}{*}{CP009494             } \\ 
&str. MC2 155 \\
\hline

\multirow{2}{*}{Genome 6}&Mycobacterium smegmatis& \multirow{2}{*}{7,092,137} &\multirow{2}{*}{NZ\_KI421511}\\   &MKD8& &\\
\hline

Genome 7& Bradyrhizobium WSM471& 7,784,016  &NZ\_CM001442\\\hline

\multirow{2}{*}{Genome 8}&Amycolatopsis thermoflava&  \multirow{2}{*}{8,677,910} &\multirow{2}{*}{NZ\_CM001442}\\   &N1165 & \\
\hline

\multirow{2}{*}{Genome 9}&Bacillus thuringiensis&   \multirow{2}{*}{6,026,843} &\multirow{2}{*}{ NZ\_CM000747 }\\    &Bt407&\\
\hline

\multirow{2}{*}{Genome 10}&Bacillus thuringiensis& \multirow{2}{*}{6,323,123 } &\multirow{2}{*}{ NZ\_CM000748}\\    &serovar thuringiensis& \\
\hline

\multirow{2}{*}{Genome 11}&Pseudomonas aeruginosa& \multirow{2}{*}{6,801,503  } &\multirow{2}{*}{NZ\_AFXI01000001}\\    &9BR&\\
\hline

\multirow{2}{*}{Genome 12}&Bacillus thuringiensis&  \multirow{2}{*}{6,260,142  } &\multirow{2}{*}{ NZ\_CM000753}\\    &serovar berliner ATCC &\\
\hline

\multirow{2}{*}{Genome 13}&Bacillus thuringiensis& \multirow{2}{*}{6,037,513 } &\multirow{2}{*}{ NZ\_CM000750 }\\    &serovar pakistani&\\
\hline

\multirow{2}{*}{Genome 14}&Pseudomonas aeruginosa& \multirow{2}{*}{6,591,121  } &\multirow{2}{*}{ CP006982  }\\    &LES400&\\
\hline
\bottomrule
\end{tabular}
\end{center}
\end{table} 
\begin{table}
\small
\label{datas}
\caption{GABM for comparison for cost and number of bits of different approaches with different datasets}
\begin{center}
\begin{tabular}{c | c c | c c | c c}
\toprule
$\textbf{Data sets}$ & \multicolumn{2}{c}{Huffman Algorithm} & \multicolumn{2}{c}{CCA }& \multicolumn{2}{c}{OCCA}\\\hline
 & $\textbf{Cost}$	& $\textbf{Bits}$ &	$\textbf{Cost}$&	$\textbf{Bits	}$&$\textbf{Cost}$&	$\textbf{Bits}$
\\\hline
Genome 1&76787151  &	37256819	&67416213 & 41332479	& 67416213 & 40503061 \\\hline
Genome 2& 100425402 &  	48778740	& 88430665	& 54662727	& 88430665&  53341739 \\\hline
Genome 3&75940155  	&36860555	&66745619	&40927187 &	66745619 & 40182241\\\hline
Genome 4&103552729  	&50047265	&90821835&	56147481&	90821835  & 54319891\\\hline
Genome 5&82234926  &	39772838	&71963876	&44254676 &	71963876 & 43392820\\\hline
Genome 6&83454842  	&40370894	&73038795	&44864009	&73038795 & 44003571\\\hline
Genome 7&92539488  &	44977876&	81416359&	49790957	&81416359 & 49127335\\\hline
Genome 8&99613856 & 	48183688	&87102639&	53817307	&87102639&	52858893\\\hline
Genome 9&71876739 &	34874617&	62998800 &	38469200&	62998800 8&	37831964\\\hline
Genome 10&75324432 & 	36576034	&66084958&	40409906	&66084958&	39739522\\\hline
Genome 11&80766360  &	39092450&	70620666	&43119140&	70620666&	42359156 \\\hline
Genome 12&74560825  &	36179579&	65359604	&39926210	&65359604&	39221148\\\hline
Genome 13&71562941 & 	34737083	&62758225 &	38299629&	62758225 &	37809443\\\hline
Genome 14&78261299 & 	37843793&	68354090&	41816580&	68354090&	 41050420\\\hline
\bottomrule
\end{tabular}
\end{center}
\end{table}

\begin{table}
\small
\label{datas}
\caption{influence of penalty on bit minimization}
\begin{center}
\begin{tabular}{c c c c }
\toprule
$\textbf{Data sets}$ & $\textbf{Cost}$	& $\textbf{Bits}$ &		\textbf{$\lambda (\%)$} \\\hline
Mycobacterium smegmatis &	 70760174 &37809636& 4\\\hline
Amycolatopsis benzoatilytica &	92010490& 49654132& 3\\\hline
Mycobacterium rhodesiae&	69421783& 37563017 &3\\\hline
Streptomyces bottropensis&	96294638& 50330594& 5\\\hline
Mycobacterium smegmatis. MC2&	74855668 &40380464& 5\\\hline
Mycobacterium smegmatis MKD8&	76738330& 40782100 &4\\\hline
Bradyrhizobium WSM471&	84667949 &45961149& 3\\\hline
Amycolatopsis thermoflava&	92416243 &48336727& 5\\\hline
Bacillus thuringiensis serovar thuringiensis&	66145783& 35330943& 4\\\hline
Bacillus thuringiensis Bt407&	68751359& 37306497 &3\\\hline
Pseudomonas aeruginosa 9BR&	72737300 &40264790& 2\\\hline
Bacillus thuringiensis serovar berliner&	67981988 &36770118 &3\\\hline
Bacillus thuringiensis serovar pakistani&	65896779& 35069999 &4\\\hline
Pseudomonas aeruginosa LES400&	71762305 &38210997 &4\\\hline
\bottomrule
\end{tabular}
\end{center}
\end{table}
\begin{figure}[h]
\begin{center}
\includegraphics{Images/Drawing4.jpg}
\caption{Convergence of Optimised cost considering Huffman algorithm}
\end{center}
\label{Fig5}
\end{figure}
The table 3 present the results founded by the basic Huffman code, cost considering algorithm and optimised cost considering algorithm. The result show that the number of bits of Huffman algorithm is the minimum number among the other algorithm but the cost is  very high. The cost considering algorithm improve the quality of the generated codes in terms of cost but the number of bits. The optimised cost considering algorithm try to find the best allocation of codes to frequencies while the cost constraint respected. The table 4 present the best founded number of bits with different penalty, for each genome we find the max number of useful penalty, after this value, increasing the penalty are in-useful (See figure 3), the number of bits achieve the minimum number but the cost stop decreasing (point 1 in figure 3) and this number of bits stabilized while the cost still increasing until it stabilized also (point 2 figure 3), after this point the cost and number of bits are stabilized.

>>>>

%% main text
\section{Introduction}
\label{sec1}
In the recent years, application of battery-powered portable devices, e.g. laptop computers and mobile phones has increased significantly. Proper representation of digital data and their transmission efficiency has become a primary concern for digital community because it affects the performance, reliability, and the cost of computation in both portable and non-portable devices. CMOS technologies were developed in order to reduce the power consumption both in data processing and transmission. In order to increase transmission speed and reduce transmission cost, parallel data transmission methods are widely used. However, parallel transmission is limited to short distance communications, e.g. locally connected devices, internal buses. Ruling out the possible availability of parallel transmission links over long distance, we are left with its serial alternative only. If we attempt to transfer big files, e.g. DNA sequences, over a serial transmission link then it would take a significant amount of time. However, we cannot overlook this problem because at present parallel processing is widely used to increase throughput and in parallel processing architecture, processing units are usually distributed in different physical locations and task sharing is a must in such architecture.    
.
Data encoding techniques came into action to improve the data transmission efficiency over serial communication medium by compressing data before transmitting. Efficiency can be measured in terms of incurred cost, required storage space, consumed power, time spent and likewise. Data must be encoded to meet the purposes like: unambiguous retrieval of information, efficient storage, efficient transmission and etc. Let a message consist of sequences of characters taken from an alphabet $\Sigma$, where  $\alpha_1,\alpha_2,\alpha_3\ldots,\alpha_r$ are the elements that represent the characters in the source $\Sigma$. The length of $\alpha_i$ represents its cost or transmission time, i.e., $c\left(\alpha_i\right)= length(\alpha_i)$. A codeword $w_i$ is a string of characters in $\Sigma$, i.e., $w_i\in\Sigma^{+}$. If a codeword is $w_i=\alpha_{i1},\alpha_{i2},\ldots,\alpha_{in}$, then the length or cost of the codeword is the sum of the lengths of its constituent elements:

@@

\section{Results And Discussion}
\label{sec4}
The approach has been evaluated with different real genomic biological data, these genomes were downloaded from a recent version of The National Center for Biotechnology Information (NCBI) available on $(http://www.ncbi.nlm.nih.gov)$. We focused on the sequences alone, ignoring any header and any other exogenous information. In table () the different data set are described with the size of each of them and the references on the biological data bank.
\begin{table}
\small
\label{datas}
\caption{Dataset description}
\begin{center}
\begin{tabular}{c  c  c c}
\toprule
$\textbf{Data sets}$ &$\textbf{Name}$ &	$\textbf{Size (MegaOctet)}$ &	$\textbf{Reference}$ \\\hline
Genome 1& Mycobacterium smegmatis &  6,66 & CP009496  \\\hline

Genome 2& Amycolatopsis benzoatilytica & 8,30 & NZ\_KB912942 \\\hline

\multirow{2}{*}{Genome 3}&Mycobacterium rhodesiae& \multirow{2}{*}{6,11} & \multirow{2} {*}{CP003169}\\ 
&NBB3& &\\
\hline

\multirow{2}{*}{Genome 4 }&Streptomyces bottropensis& \multirow{2}{*}{8,54} &
\multirow{2}{*}{NZ\_KB911581} \\ 
& ATCC 25435 \\
\hline

\multirow{2}{*}{Genome 5}&Mycobacterium smegmatis& \multirow{2}{*}{ 6,66} &\multirow{2}{*}{CP009494             } \\ 
&str. MC2 155 \\
\hline

\multirow{2}{*}{Genome 6}&Mycobacterium smegmatis& \multirow{2}{*}{6,76} &\multirow{2}{*}{NZ\_KI421511}\\   &MKD8& &\\
\hline

Genome 7& Bradyrhizobium WSM471&  7,42 &NZ\_CM001442\\\hline

\multirow{2}{*}{Genome 8}&Amycolatopsis thermoflava&  \multirow{2}{*}{8,27} &\multirow{2}{*}{NZ\_CM001442}\\   &N1165 & \\
\hline

\multirow{2}{*}{Genome 9}&Bacillus thuringiensis&   \multirow{2}{*}{5,74} &\multirow{2}{*}{ NZ\_CM000747 }\\    &Bt407&\\
\hline

\multirow{2}{*}{Genome 10}&Bacillus thuringiensis& \multirow{2}{*}{6,03} &\multirow{2}{*}{ NZ\_CM000748}\\    &serovar thuringiensis& \\
\hline

\multirow{2}{*}{Genome 11}&Pseudomonas aeruginosa& \multirow{2}{*}{6,48} &\multirow{2}{*}{NZ\_AFXI01000001}\\    &9BR&\\
\hline

\multirow{2}{*}{Genome 12}&Bacillus thuringiensis&  \multirow{2}{*}{5,97} &\multirow{2}{*}{ NZ\_CM000753}\\    &serovar berliner ATCC &\\
\hline

\multirow{2}{*}{Genome 13}&Bacillus thuringiensis& \multirow{2}{*}{5,75} &\multirow{2}{*}{ NZ\_CM000750 }\\    &serovar pakistani&\\
\hline

\multirow{2}{*}{Genome 14}&Pseudomonas aeruginosa& \multirow{2}{*}{6,28} &\multirow{2}{*}{CP006982}\\    &LES400&\\
\hline

\multirow{2}{*}{Genome 15}& Mus musculus & \multirow{2}{*}{25,58} &\multirow{2}{*}{GL456087}\\  &chromosome 1&\\
\hline

\multirow{2}{*}{Genome 16}& Danio rerio & \multirow{2}{*}{56,14} &\multirow{2}{*}{CM002885}\\ & chromosome 1 &\\
\hline
\multirow{2}{*}{Genome 17}&Homo sapiens & \multirow{2}{*}{76,64 } &\multirow{2}{*}{  CM000680  }\\    & chromosome 18 &\\
\hline
\multirow{2}{*}{Genome 18}&Homo sapiens & \multirow{2}{*}{74,92} &\multirow{2}{*}{CM000684   }\\ & chromosome 22&\\
\hline
\bottomrule
\end{tabular}
\end{center}
\end{table} 
\begin{table}
\small
\label{datas}
\caption{GABM for comparison for cost and number of bits of different approaches with different datasets}
\begin{center}
\begin{tabular}{c | c c | c c | c c}
\toprule
$\textbf{Data sets}$ & \multicolumn{2}{c}{Huffman Algorithm} & \multicolumn{2}{c}{CCA }& \multicolumn{2}{c}{OCCA}\\\hline
 & $\textbf{Cost}$	& $\textbf{Bits}$ &	$\textbf{Cost}$&	$\textbf{Bits	}$&$\textbf{Cost}$&	$\textbf{Bits}$
\\\hline
Genome 1& 76787151  &	4,44	&67416213   & 4,92	& 67416213 & 4,82 \\\hline
Genome 2& 100425402 &  	5,81	& 88430665	& 6,51	& 88430665&  6,35 \\\hline
Genome 3& 75940155  &	4,39    &66745619	& 4,87&	66745619 & 4,79 \\\hline
Genome 4& 103552729 &	5,96    &90821835&	6,69 &	90821835  & 6,47\\\hline
Genome 5& 82234926  &	4,74	&71963876	& 5,27 &	71963876 &5,17 \\\hline
Genome 6& 83454842  &   4,81	&73038795	& 5,34	&73038795 & 5,24\\\hline
Genome 7& 92539488  &	5,36    &81416359   &5,93	&81416359 & 5,85\\\hline
Genome 8& 99613856  & 	5,74	&87102639   &6,41	&87102639&	6,30 \\\hline
Genome 9& 71876739  &	4,15    &62998800  &4,58 &	62998800 8&	5,50 \\\hline
Genome 10& 75324432 & 	4,36	&66084958  &4,81	&66084958&	4,73 \\\hline
Genome 11& 80766360 &	4,66    &70620666  &5,41&	70620666&	5,04 \\\hline
Genome 12& 74560825 &	4,31    &65359604  &4,75	&65359604&	4,67\\\hline
Genome 13& 71562941 & 	4,14	&62758225  &4,56	&	62758225 &	4,50\\\hline
Genome 14& 78261299 & 	4,51    &68354090  &4,98&	68354090& 4,89  \\\hline
Genome 15& 324439242& 	18,77    &286008420 & 20,66	 &	288866525 &	19,94 \\\hline
Genome 16& 703734840&  40,77    &618859291 & 45,08  &	625042576 &	43,18 \\\hline
Genome 17& 901032667&  52,08    &791840455 &	57,36 &	799757406 &	55,12 \\\hline
Genome 18& 983434816&  56,98    &867299889 &	62,42 &	875969422 & 60,70 \\\hline

\bottomrule
\end{tabular}
\end{center}
\end{table}

\begin{table}
\small
\label{datas}
\caption{influence of penalty on bit minimization}
\begin{center}
\begin{tabular}{c c c c c }
\toprule
$\textbf{Data sets}$ & $\textbf{Cost}$	& $\textbf{Bits}$ &		\textbf{$\lambda (\%)$} &		\textbf{Compression ratio $\%$} \\\hline
Genome 1 &	 70760174 &4,50& 4 & 32,72\\\hline
Genome 2 &	92010490& 5,91 & 3& 28,79\\\hline
Genome 3&	69421783& 4,47 &3& 26,84\\\hline
Genome 4&	96294638& 5,99 & 5& 29,85\\\hline
Genome 5&	74855668 &4,81& 5& 27,77\\\hline
Genome 6&	76738330& 4,86  &4& 28,10\\\hline
Genome 7&	84667949 &5,47& 3& 26,28\\\hline
Genome 8&	92416243 &5,76& 5& 30,35\\\hline
Genome 9&	66145783&4,21 & 4& 26,65\\\hline
Genome 10&	68751359& 4,44 &3& 26,63\\\hline
Genome 11&	72737300 &4,79& 2&26,08 \\\hline
Genome 12&	67981988 & 4,38&3& 26,63\\\hline
Genome 13&	65896779& 4,18 &4& 27,30\\\hline
Genome 14&	71762305 &4,55 &4&27,54 \\\hline
Genome 15&297188002 &19,31 & 3& 24,51\\\hline
Genome 16&643785655 & 41,68 & 3& 25,75 \\\hline
Genome 17&823506882 & 53,05 & 3& 30,78\\\hline
Genome 18&901515198 &58,99 & 3& 21,26\\\hline
\bottomrule
\end{tabular}
\end{center}
\end{table}
\begin{figure}[h]
\begin{center}
\includegraphics{Images/Drawing4.jpg}
\caption{Convergence of Optimised cost considering Huffman algorithm}
\end{center}
\label{Fig5}
\end{figure}
The table 3 present the results founded by the basic Huffman code, cost considering algorithm and optimised cost considering algorithm. The result show that the number of bits of Huffman algorithm is the minimum number among the other algorithm but the cost is  very high. The cost considering algorithm improve the quality of the generated codes in terms of cost but the number of bits. The optimised cost considering algorithm try to find the best allocation of codes to frequencies while the cost constraint respected. The table 4 present the best founded number of bits with different penalty, for each genome we find the max number of useful penalty, after this value, increasing the penalty are in-useful (See figure 3), the number of bits achieve the minimum number but the cost stop decreasing (point 1 in figure 3) and this number of bits stabilized while the cost still increasing until it stabilized also (point 2 figure 3), after this point the cost and number of bits are stabilized.
